Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments

### 2022-04-09 doc/peakcharacterization_manuscript.Rmd
new version of p-value comparison by peak source
using a t-test for examination of the differences we see, and putting the summary values directly on the sina-plot so we don't actually need a table

::: {custom-style="MDPI_3.1_text"}
For large datasets, we expect that more correct peak intensities will result in better agreement between sample normalized peak intensities within a sample class.
One way to test this is to compare p-values generated from the different intensities.
:::

lung_compare %>%
  ggplot(aes(x = log_p.ref,
             y = log_p.test)) +
  geom_point() +
  geom_abline(slope = 1, color = "red") +
  facet_wrap(~ id.test, ncol = 1) +
  labs(x = "-1 * Log10(P-Value) Raw Scan-Centric",
       y = "-1 * Log10(P-Value) Other")
```
:::

::: {custom-style="MDPI_5.1_figure_caption"}
`r figure_count$label_text("p_value_comparison")`.
Log-p-values generated by comparing non-cancer and cancer sample IMFs using peak intensities from different methods.
:::

::: {custom-style="MDPI_4.1_table_caption"}
`r table_count$label_text("p_value_comparison")`.
Binomial test results on log-p-value differences to p-values generated by raw scan-centric IMF intensities.
P-values were adjusted using the Bonferroni method.
:::

::: {custom-style="MDPI_4.2_table_body"}
```{r p_value_comparison_table}
tar_load(lung_binomial)
binomial_order = data.frame("comparison" = c("corrected", "xcalibur", "msnbase"))
lung_binomial = dplyr::left_join(binomial_order,
                                 lung_binomial,
                                 by = "comparison")
lung_binomial %>%
  dplyr::transmute(Comparison = comparison,
                `N-Positive` = n_positive,
                `N-Negative` = n_negative,
                `P-Value` = sprintf("%.02e", p.value),
                `Adj-P-Value` = sprintf("%.02e", p.adjust)) %>%
  flextable() %>%
  align(j = c(4, 5), align = "right", part = "all") %>%
  autofit()
```
:::

>>>>

::: {custom-style="MDPI_3.1_text"}
For large datasets, we expect that more correct peak intensities will result in better agreement between sample normalized peak intensities within a sample class.
One way to test this is to compare p-values generated from the different intensities.
`r figure_count$label_text("p_value_comparison")` compares the -1 * Log10(p-values) from `r nrow(compare_split$Corrected)` (corrected and XCalibur) or `r nrow(compare_split$MSnbase)` (MSnbase) isotope resolved molecular formulas (IMFs) assigned by SMIRFE for the scan-centric peaks and then peaks matched and intensities extracted from the other methods.
Although all of the p-vaues are somewhat (and statistically) different from those reported by the raw intensities, there are some interesting patterns of differences.
MSnbase generated p-values show the widest distribution of differences, as well as the smallest number of peaks that are present in 50% of both the cancer and non-cancer samples.
This echoes the patterns of low overlapping assigned peaks observed in both ECF samples, where many of the scan-centric amino-acid peaks only found one match from the MSnbase peaks.
Surprisingly, the corrected intensities generated very different p-values compared to the raw intensities, and XCalibur showed the most agreement with the raw p-values.
We tested the statistical significance of these differences using a t-test of the actual difference in log-p-values (shown in `r figure_count$label_text("p_value_comparison")`D).
:::

::: {custom-style="MDPI_5.2_figure"}
```{r p_value_comparison, dn_id = figure_count, fig.width = 8, fig.height = 8}
tar_load(lung_compare_ttest)

compare_plots = purrr::imap(compare_split, function(in_compare, id){
  out_plot = in_compare %>%
    ggplot(aes(x = log_p.ref, y = log_p.test)) +
    geom_abline(slope = 1, color = "red") +
    geom_point(alpha = 0.5) +
    coord_equal() +
    labs(x = "Log-P-Value Raw",
         y = paste0("Log-P-Value ", id))
  out_plot
})

long_ttest = lung_compare_ttest %>%
  dplyr::select(p.adjust, comparison, estimate, conf.low, conf.high) %>%
  tidyr::pivot_longer(cols = !comparison, names_to = "measure", values_to = "value") %>%
  dplyr::mutate(compare_id = dplyr::case_when(
    comparison %in% "corrected" ~ "Corrected",
    comparison %in% "xcalibur" ~ "XCalibur",
    comparison %in% "msnbase" ~ "MSnbase"
  ))

long_pvalue = long_ttest %>%
  dplyr::filter(measure %in% "p.adjust")

long_measure = long_ttest %>%
  dplyr::filter(!(measure %in% "p.adjust"))

compare_sina = lung_compare %>%
  ggplot(aes(x = compare_id, y = p_diff)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_sina(alpha = 0.5) +
  geom_point(data = long_measure, aes(x = compare_id, y = value), color = "green") +
  geom_text(data = long_pvalue, aes(x = compare_id, y = -5, label = format(value, digits = 2))) +
  labs(x = "", y = "Log-P-Value Difference")

lung_plot = (compare_plots$Corrected | compare_plots$MSnbase) / (compare_plots$XCalibur | compare_sina) +
  plot_annotation(tag_levels = "A")

lung_plot
```
:::

::: {custom-style="MDPI_5.1_figure_caption"}
`r figure_count$label_text("p_value_comparison")`.
**A - C**: Log-p-values generated by comparing non-cancer and cancer sample IMFs using peak intensities from different methods.
Red line denotes perfect agreement.
**D**: Sina plot of differences in the log-p-values generated by different methods compared to the raw scan-centric log-p-values.
Also shown are the Bonferroni adjusted p-values from a t-test of the log-p-value differences for each method.
Green points denote the high, mean, and low-confidence limits reported from the t-test.
:::
